---
title: "referenceExtract"
subtitle: "A vignette using California's Groundwater Sustainability Plan documents"
output: html_document
---
<!----
These were in the yaml but the html won't render to the vignette folder and that is annoying to me, so instead i run this: rmarkdown::render('vignettes/sgma.Rmd', output_dir = 'vignettes')
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sgma}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
--->
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = F
)
```

The referenceExtract tool from govscienceuseR is designed to take unstructured PDF documents, feed them through the anystyle.io reference extraction software, and return tagged reference data. We walk through these steps using California's Groundwater Sustainability Plan PDF documents.  

First, be sure to download the referenceExtract package from github (`devtools::install_github("govscienceuseR/referenceExtract")` and load it in, as below. Also load the packages listed below.  

```{r setup, results = F, warning = F, message = F}
library(referenceExtract)
packages = c('data.table','pdftools','pbapply','stringr', 'jsonlite', 
             'purrr', 'magrittr', 'dplyr', 'tidyr', 'tools')
sapply(packages, require, character.only = T)
```

Now let's take a look at PDFs we will be analyzing. The PDFs have been downloaded all into one folder, listed below.  

```{r}
list.files("~/Box/reference_classifier/documents_gsp/")
```

We will take these files and walk through the three steps of the referenceExtract package: extracting with `reference_extract`, compiling with `reference_compile`, and cleaning with `reference_clean`.  

The `reference_extract` function below reads in every PDF in the document directory (doc_dir), 'documents_gsp/' folder, and runs them through [anystyle.io](https://anystyle.io/). Anystyle extracts probable references and exports them to the reference directory (ref_dir), 'reference_extracts_gsp' as JSON files.. We specify no layout to the documents.  

```{r, eval = F}
reference_extract(doc_dir = '~/Box/reference_classifier/documents_gsp/', 
                 ref_dir = '~/Box/reference_classifier/reference_extracts_gsp', 
                 layout = "none")
```

After extractions we a JSON file for each PDF document.  

```{r}
list.files('~/Box/reference_classifier/reference_extracts_gsp')
```

Next, the `reference_compile` function will transform those JSONs to tabular data and compile them all into one data frame, adding the file name as an identifier.  

```{r, eval = F, warning = F}
dt <- reference_compile('~/Box/reference_classifier/reference_extracts_gsp')
```

```{r, echo = F}
# Right now reference_compile is bugging when only in the package, so for now overwriting it. Output from when reference_compile runs outside of package
dt <- readRDS("~/Box/reference_classifier/data/gsp_references.RDS")
```

We can first look at this raw output from Anystyle. It organizes potential references into the following categories:  

```{r}
colnames(dt)
```

But the predicted references are a bit messy. Authors are nested into a matrix, some rows have listed values, etc.  

```{r}
head(as_tibble(select(dt, title, author, date, publisher, `container-title`, doi)), n = 10)
```

To try to clean up these potential references, the `reference_clean` function goes through a series of steps. For each column the function unlists the data and filters out unlikely candidates for that column. For instance, if a number listed in the date column does not match any reasonable date format or expectation, it is removed. If a string in the URL column actually resembles a DOI, it is moved to that column. And so on. 


```{r, eval = F}
cleaned_dt <- reference_clean(dt)
```

```{r, echo = F}
# Output from when reference_clean runs
cleaned_dt <- fread("~/Box/reference_classifier/data/gsp_references_clean.csv")
```

The product is an unlisted and filtered data frame, as below.  

```{r}
head(as_tibble(select(cleaned_dt, title, author, year, publisher, `container`, doi)), n = 10)
```

Let's save this file for use in our next step in the workflow, the [referenceClassify](https://github.com/govscienceuseR/referenceClassify) package. 
```{r, eval = F}
fwrite(cleaned_dt, "~/Box/reference_classifier/data/gsp_references_clean.csv")
```


